---
title: "MSDA Week2 Assignment"
author: "Abhishek Kumar"
date: "July 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment


In probability, "n choose r" means n! / (( n - r)! * r!).
For example, "5 choose 3" evaluates to 10.
R has built in functions for choose() and factorial(). Your task is to write your own variation of the choose() function, that does not use either of these functions.
Your code should adhere to the style guidelines provided by Hadley Wickham in this week's reading.
You need to create your code and some test code, and deliver your code, results, and documentation in a single R Markdown file.

## Solution

We first need to design factorial function without using existing factorial function.

#### Custom Factorial Function

By definition factorial of any positive integer $n$  $(n >=0)$ is defined as 
$n! = (n) * (n-1) *(n-2) ... 3 * 2 * 1$.

```{r}
custom.factorial <- function(x){
  if(!is.integer(x) | x < 0) {
    paste("Factorial is defined for postive integeres only")
  }
    if( (x==0 || x==1) & x >=0) {
    1
  }
  else {
    y <- 1
    for (i in 1:x) {
      y <- y*i
    }
    y
  }    
}
```

Below are some test results for custom.factorial() function.

```{r}
m <- custom.factorial(5)
m
n <- custom.factorial(10)
n
```

#### Custom Choose Function
The number of ways of picking k unordered outcomes from n possibilities. Also known as the binomial coefficient or choice number and read "n choose k" is defined as 

${n \choose k} = n! / ((n-k)! * k!)$    whenever $k <= n$

Now we can design the custom choose function using custom factorial function.

```{r}
custom.choose <- function(n,r) {
  custom.factorial(n) / (custom.factorial(n-r) * custom.factorial(r))
}
```

Below are some test results of custom.choose() function

```{r}
a <- custom.choose(5,2)
a
```


