---
title: "MSDA Final Project"
subtitle : "Analyze Airport Departure Data of Houston Airport"
author: "Abhishek Kumar"
date: "August 8, 2016"
output:
  ioslides_presentation:
    fullscreen : true
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Agenda

 - Understand the requirement of Final Project Task.
 - Understand Data
 - Define Analytics Requirement.
 - Identify Key Performance Indicators (KPIs)
 - Analyze the Data
 
## MSDA Final Project

**Task**

This project is where you show off your ability to 

- use R packages
- change the shape of data in a data frame
- provide basic summary statistics and graphics as part of your exploratory   data analysis.

**Instructions**

- You should start by installing and loading the R package hflights, which is available on CRAN and described here: https://cran.r-project.org/web/packages/hflights/hflights.pdf
- You should perform some basic transformations on the data frame, such as including only a subset of columns, created one or more new derived columns (such as flight date), and/or filtering out rows.
- You should then create basic summary statistics. You should create at least one interesting question for analysis, and answer the question using two or more graphics in R.



## Understand Data
**hflights** dataset contains all flights *departing* from Houston airports IAH (George Bush Intercontinental) and HOU (Houston Hobby) in *2011*. The data comes from the Research and Innovation Technology Administration at the Bureau of Transporation statistics(BTS) provided by author Hadley Wickham.
It is a data frame with 227,496 rows and 21 columns.

```{r}
library(hflights)
library(ggplot2)
library(plyr)
head(hflights)
```


<div class="notes">
**Variables**  

- Year, Month, DayofMonth: date of departure 
- DayOfWeek: day of week of departure (useful for removing weekend effects) 
- DepTime, ArrTime: departure and arrival times (in local time, hhmm) 
- UniqueCarrier: unique abbreviation for a carrier 
- FlightNum: flight number 
- TailNum: airplane tail number 
- ActualElapsedTime: elapsed time of flight, in minutes 
- AirTime: flight time, in minutes 
- ArrDelay, DepDelay: arrival and departure delays, in minutes 
- Origin, Dest origin and destination airport codes 
- Distance: distance of flight, in miles 
- TaxiIn, TaxiOut: taxi in and out times in minutes 
- Cancelled: cancelled indicator: 1 = Yes, 0 = No 
- CancellationCode: reasonforcancellation: A=carrier,B=weather,C=nationalairsystem, D = security 
- Diverted: diverted indicator: 1 = Yes, 0 = No

</div>

## Define Analytics Requiremnet
According to BTS,  A flight is considered on-time when it arrives less than 15 minutes after its published arrival time.

Based on above assumption and existing Dataset below are the list of Analytics Requiremnet.


> * **Overall Summary on Airport Data.**
> * **Which airlines are good for On Time Arrival at Scheduled Location.**
> * **Which airlines are good for On Time Departure at SCheduled Location.**
> * **Which attributes are responsible for cancelattion of Major flights.**



## Key Performing Indicator & Metrics

> * **Arrival Timeliness (%)** 
    - On Time Arrival Count : - Count of flights which are arrived within                                  15 minutes of scheduled Arrival Time. 
    - Total Count :- Count of all flights which are not cancelled.
    - $$Arrival Timeliness = On Time Arrival Count / Total Count *100$$ 

> * **Departure Timeliness (%)** 
    - On Time Departure Count : - Count of flights which are arrived within                                  15 minutes of scheduled Arrival Time. 
    - Total Count :- Count of all flights which are not cancelled.
    - $$Departure Timeliness = On Time Departure Count / Total Count *100$$ 
> * **Number of Flights Departed on Scheduled Route**

> * **Number of Flights Departed on Diverted Route**

> * **Number of Flights Cancelled**



## Basic Summary {.smaller .columns-2}


```{r, warning=FALSE, out.width =  '400px', out.height= '500px'}
a4 <- unique(hflights[,c(14:15)])
a4$Origin[a4$Origin == 'IAH'] <- "George Bush Intercontinental"
a4$Origin[a4$Origin == 'HOU'] <- "Houston Hobby"
g1 <- ggplot(a4,aes(Origin))
g1 + geom_bar(aes(fill = Origin)) + xlab('Origin Airport') + ylab('Number of Destination') + ggtitle("Summary of Airline connected to Houston Airports")
```




```{r, warning=FALSE, out.width =  '400px', out.height= '500px'}
b1 <- hflights[,c(2,7:9,12:15,19:21)]
b1$type <- NA
b1$type[b1$Cancelled == 1] <- 'Cancelled'
b1$type[b1$Diverted == 1] <- 'Diverted'
b1$type[b1$Cancelled == 0 & b1$Diverted == 0] <- 'OnRoute'
b1$MonthName <- month.abb[b1$Month]
b1$MonthName <- factor(b1$MonthName, levels = b1$MonthName[order(b1$Month)])
#b2 <- ddply(b1,c('Month','MonthName','type'),nrow)
#head(b2)
#monthly_status <- ggplot(b2,aes(Month,V1))
#monthly_status + geom_line(aes(color = type))
month_summary <- ggplot(b1,aes(MonthName)) 
month_summary + geom_bar(width = 0.25,aes(fill = MonthName)) +  facet_grid(type ~ ., scale="free_y") + xlab('Month') + ylab('Number of Flights') + ggtitle("Monthly Status Report") + labs(fill = "Month Name")
```



## Timeliness {.smaller .columns-2}

```{r,,eval=TRUE, warning=FALSE, out.width =  '400px', out.height= '500px'}
library(plyr)
a7 <- hflights[,c(2,5,6,7,10:15)]
a7$OnTimeArr <- ifelse(a7$ArrDelay <=15, 1,0)
a7$OnTimeDept <- ifelse(a7$DepDelay <=15, 1,0)
s1 <- ddply(a7,'UniqueCarrier',function(x) c(count=nrow(x),sum=sum(x$OnTimeDept, na.rm = TRUE), percent = sum(x$OnTimeDept, na.rm = TRUE)*100/nrow(x) ))
#head(s1)
s1$UniqueCarrier <- factor(s1$UniqueCarrier, levels = s1$UniqueCarrier[order(s1$percent)])
g2 <- ggplot(s1,aes(UniqueCarrier,percent,group =1))
g2 + geom_bar(stat = "identity",aes(fill = UniqueCarrier) ) + geom_line() + xlab('Airline Code') + ylab('Arrival Timeliness(%)') + ggtitle("Airline Arrival Timeliness") + labs(fill = "Airline Code")
```



```{r ,eval=TRUE, warning=FALSE, out.width =  '400px', out.height= '500px'}
s2 <- ddply(a7,'UniqueCarrier',function(x) c(count=nrow(x),sum=sum(x$OnTimeArr, na.rm = TRUE), percent = sum(x$OnTimeArr, na.rm = TRUE)*100/nrow(x) ))
s2$UniqueCarrier <- factor(s2$UniqueCarrier, levels = s2$UniqueCarrier[order(s2$percent)])
g2 <- ggplot(s2,aes(UniqueCarrier,percent,group =1))
g2 + geom_bar(stat = "identity",aes(fill = UniqueCarrier) ) + geom_line() + xlab('Airline Code') + ylab('Departure Timeliness(%)') + ggtitle("Airline Departure Timeliness") + labs(fill = "Airline Code")
```

## Cancellation {.smaller .columns-2}

```{r, warning = FALSE, out.width =  '400px', out.height= '500px'}
a9 <- subset(hflights,hflights$Cancelled == 1)
a9$CancellationCode[a9$CancellationCode == 'A'] <- 'Carrier'
a9$CancellationCode[a9$CancellationCode == 'B'] <- 'Weather'
a9$CancellationCode[a9$CancellationCode == 'C'] <- 'National Air System'
a9$CancellationCode[a9$CancellationCode == 'D'] <- 'Security'
a9$Origin[a9$Origin == 'IAH'] <- "George Bush Intercontinental"
a9$Origin[a9$Origin == 'HOU'] <- "Houston Hobby"
a9$MonthName <- month.abb[a9$Month]
a9$MonthName <- factor(a9$MonthName, levels = a9$MonthName[order(a9$Month)])
b2 <- ddply(a9,c('Month','MonthName','CancellationCode'),nrow)
#head(b2)
#b2$MonthName <- factor(b2$MonthName, levels = b2$MonthName[order(b2$Month)])
#monthly_status <- ggplot(b2,aes(MonthName,V1))
#monthly_status + geom_line(aes(color = CancellationCode))
monthly_cancellation <- ggplot(a9,aes(MonthName)) 
monthly_cancellation + geom_bar(width = 0.25,aes(fill = MonthName)) +  facet_grid(CancellationCode ~ ., scale="free_y") + xlab('Month') + ylab('Number of Flights Cancelled') + ggtitle("Monthly Status Report\nfor Cancelled Flights") + labs(fill = "Month Name")

```


```{r,warning = FALSE, out.width =  '400px'}
#library(plotrix)

pie2 <- ggplot(a9, aes(x = factor(1), fill = factor(CancellationCode))) +
  geom_bar(position = "fill",width = 1) + facet_grid(facets=. ~ Origin)
pie2 + coord_polar(theta = "y") + xlab('') +  ylab('') + 
  labs(fill='Cancellation Code') + scale_fill_brewer(palette="Dark2") +   ggtitle("Factors Contributing to\nFlight Cancellation")
```


## References

- `hflights` package by Hadley Wickham 
  https://cran.r-project.org/web/packages/hflights/hflights.pdf
- Research and Innovation Technology Administration at the Bureau of       Transporation statistics
http://www.transtats.bts.gov/DatabaseInfo.asp?DB_ID=120&Link=0
